### 2020-07-25 tex/results.tex
fix bit about convergence with sharp edges

%!TEX root = ClementiBarba2020.tex

\subsection{Replication of results from Rockstuhl, et al., 2005}

%%% Rockstuhl work summary%%%%
The work of Rockstuhl et al.\cite{rockstuhl2005} studies the phonon-polariton response of silicon carbide (SiC)
nanoparticles using the boundary element method. They analyze different ``cylindrical particles''
made of 6H-SiC, as a function of geometrical cross-section. 
They use a two-dimensional (2D) boundary element method 
previously developed by the group \cite{rockstuhl2003}, where they extend in the third dimension to infinity
(hence the name ``cylinder'').
%%% Rockstuhl work summary%%%%

We decided to attemp to replicate one of the results presented on Fig.14 of their paper \cite{rockstuhl2005}:
the scattering cross-section of a SiC rectangular cylinder for three different aspect ratios. 
To comply with our quasistatic approach ($\lambda > d$ where $d$ is the characteristic
dimension of the geometry), we chose the particular case where a=672 nm and b=328 nm.

\subsubsection{Differences in method, mesh and dielectric data}

\paragraph{Method:} The main difference between the original simulations and ours is that they solve a 2D problem with the 
full Maxwell equations, while we solve a 3D problem with the electrostatic approximation. 
We lack any information about their code implementations, discretization schemes, or solver.  
We computed extinction cross-section (scattering plus absorption) while Rockstuhl et al.\ present only scattering cross-section. 
But, since the quantity of interest is the wavelength at which the resonance modes occur, the results can be compared.

\paragraph{Mesh:} Rockstuhl et al.\ did not provide any details regarding the discretization of the geometries or 
parameters involved in the simulations.
We performed a grid-independence study as a form of solution verification, and to ensure that we are 
minimizing errors due to discretization. Rockstuhl et al.\ used a 2D geometry (infinite third dimension), while we treat the geometry in its full 3D representation.

\paragraph{Dielectric data:} The study uses 6H-SiC as material, and the authors obtained their data from a a source that we
were not able to replicate. Instead we are using experimental data for 4h-SiC that was provided to us 
via private communications from the authors of Ellis et al.\cite{ellis2016}.

\subsubsection{Grid-independence study}

We performed a grid-independence study on a SiC cube of side $L=535$ nm submerged in air, under a 
constant electric field aligned with the $z$-axis (a similar setup to the square cylinder on Fig. 18 of 
Rockstuhl et al.\cite{rockstuhl2005}). 
Due to the nature of the geometry and its sharp edges, it was challenging to see proper convergence. 
We have completed grid-convergence analysis for this type of physics in a previous work, but using spherical geometries \cite{ClementiETal2019}. 
With that experience, and given the difficulty caused by sharp edges, we are satisfied with studying grid-independence here instead.
Figure \ref{fig:cube535} shows the grid-independence study, using meshes with  15,552 triangles (density $9.05\times10^{-4}$ per $\text{\AA}$ squared)
 to 19,200 triangles (density $1.11\times10^{-5}$ $\text{\AA}$ squared):
 the computed results show no discernible difference.

\begin{figure}
    \centering
    \includegraphics[width=0.75\textwidth]{cubeL535nm_15Kvs19K.pdf} 
    \caption{Grid-independence study for a SiC cube of side $L=535 nm$ submerged in air under a constant 
    electric field in the $z$-direction. The curves represent the extinction cross-section divided by $L^2$ 
    as a function of wavelength divided by $L$.}
    \label{fig:cube535}
 \end{figure}

It is worth noting that the extinction cross-section curve in Figure \ref{fig:cube535} has extra peaks 
compared to the results of Figure 18 of Rockstuhl et al., due to the three-dimensional effects captured in our simulation and the sharp 
edges, the latter effect also being mentioned by Rockstuhl et al. The 3D effects will be approached
in the following results, where we attempt a replication of one of the results on Figure 14 of Rockstuhl et al.

\subsubsection{Replication of Figure 14 (case a1) of Rockstuhl et al., 2005}

We chose to replicate a result of Rocksuthl et al.\ presented in Figure 14: 
the case where $a=672$ nm 
and $b=328$ nm, since these dimensions are well within the limits of the quasistatic approximation 
used in \pygbe. Rockstuhl et al.\ present the normalized scattering cross-section of a SiC rectangular 
cylinder, and they perform simulations for two different setups, sketched in Figure \ref{fig:rectangle_sketch}. In 
their Figure 14 (left), they have the wave vector (illumination) along the long 
side of the geometry, which means that the electric field is parallel to the short side of the rectangle, like in 
Figure \ref{fig:rectangle_sketch} (B). Following a similar analysis, in Figure 14 (right) of Rockstuhl et al., they have the wave 
vector (illumination) along the short side of the geometry, which means that the electric field is parallel to the 
long side of the rectangle, like in Figure \ref{fig:rectangle_sketch} (A).

Based on the results of our grid-independence study, we decided to use triangle densities in the order of $6$ to $9\times10^{-5}$
to produce the meshes for the replication of figures 14a and 14b in the work of Rockstuhl et al.\ (case $a1=672$, $b=328$). For the third 
dimension, we needed to choose a value that represents ``infinity.'' To achieve this, we studied the effects of 
elongating the cylinder to the third dimension. In Figure \ref{fig:ext_y_14}, we present the results for two different
values of the length in the third dimension, $y=1344$ nm ($2\times a$) and $y=2688$ nm ($4\times a$). We see that as we make $y$ longer, the 
intensity of some peaks decreases, due to the fact that some of the peaks are associated to this component. Therefore, 
from now on we use $y=2688$ nm. We could have increased this value more, but the
simulations then become computationally more expensive and we would still have the 3D effects in the model.

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{rockstuhl_rectangles.pdf} 
    \caption{Rockstuhl runs configurations}
    \label{fig:rectangle_sketch}
\end{figure}

\begin{figure}
    \centering
    \subfloat{\includegraphics[width=0.48\textwidth]{ext_y_14a.pdf}}
    \subfloat{\includegraphics[width=0.48\textwidth]{ext_y_14b.pdf}} 
    \caption{Effect of the elongation of the third direction ($y$) on the 
        extinction cross section of a rectangular prism of SiC of dimensions $a=672$ nm 
        and $b=328$ nm, submerged in air and under a constant electric field 
        parallel to the z-axis. The left plot corresponds to a configuration such that the electric 
        field is parallel to $b$ (configuration (A) on Figure \ref{fig:rectangle_sketch}), while the 
        right plot corresponds to a configuration such that the electric field is 
        parallel to $a$ (configuration (B) on Figure \ref{fig:rectangle_sketch}}
    \label{fig:ext_y_14}   
 \end{figure}

To generate the meshes for these simulations, we initially used the open source software Trimesh 
(\url{https://github.com/mikedh/trimesh}), but we realized that it was not producing a 
uniform mesh and that it was not possible to obtain regular triangles with the functions 
available when having a prism. To overcome this, we created our own mesh using Python scripts,
and obtained uniform meshes. We wanted to study the effect of a uniform mesh as well as the effect
of rounding the edges---Rockstuhl et al.\ mentioned rounded edges
as a factor that introduces extra peaks on the response. We were unable to control the 
roundness as a function of arc of curvature or the dimensions of the rectangular prism, so we 
decided to use the default settings on Trimesh. Figure \ref{fig:tri_reg_round_14} shows the 
results of the effect of uniformity and roundness. One can see that the second peak is not
present in the green curve, which can be attributed to the effect of the roundness. This is 
consistent with the results in Rockstuhl et al.

\begin{figure}
    \centering
    \subfloat{\includegraphics[width=0.48\textwidth]{tri_reg_round_14a.pdf}}
    \subfloat{\includegraphics[width=0.48\textwidth]{tri_reg_round_14b.pdf}}
    \caption{ Effect of uniformity and roundness on the mesh on the 
    extinction cross section of a rectangular prism of SiC of dimensions $a=672$ nm, 
    $b=328$ nm and $y=2688$ nm, submerged in air and under a constant electric field 
    parallel to the z-axis}
    \label{fig:tri_reg_round_14}
 \end{figure}

Once we have found the ``best'' possible geometry construction, we show how our results 
(green curve in Figure \ref{fig:tri_reg_round_14}) compare 
with the original results from Rockstuhl's Figure 14. We have obtained Rockstuhl's curves by using 
the WebPlotDigitizer (\url{https://apps.automeris.io/wpd/}). The replication results are 
presented in Figure \ref{fig:rep_14}: the main peaks presented in Rockstuhl et al.\ are 
successfully replicated. While we still have the presence of a third peak in our results, we conjecture 
that this is a consequence of the 3D effects in our geometry.

 \begin{figure}
    \centering
    \subfloat{\includegraphics[width=0.48\textwidth]{replication_14a.pdf}}
    \subfloat{\includegraphics[width=0.48\textwidth]{replication_14b.pdf}} 
    \caption{Replication of Figure 14 from Rockstuhl 2005. Extinction cross section of a
    rectangular prism of SiC of dimensions $a=672$ nm, $b=328$ nm and $y=2688$ nm, submerged
    in air and under a constant electric field parallel to the z-axis.}
    \label{fig:rep_14}
 \end{figure}

 \subsection{Replication of results from Ellis et al., 2016, and validation}

 %%% Ellis work summary %%%%
The work of Ellis and coworkers \cite{ellis2016} study the aspect-ratio evolution of high-order
resonant modes in localized surface phonon-polariton nanostructures. They study the
excitation of multipolar localized surface phonon polaritons (SPhP) resonances, by measuring
polarized reflectance on 4H-SiC pillars of fixed height ($H=950$ nm), fixed 
width ($W=400$ nm) and varied length ($L=400-4800$ nm) and therefore an aspect ratio
($AR=L/W=1-12$). These pillars are patterned on a squared grid with a pitch $P=L+500$ nm
to reduce coupling. In  both their simulations and experiments they excite the localized SPhP
resonances, they perform polarized reflectance measurements with the incident polarization 
oriented parallel or perpendicular to the long axis of the pillars.  
 %%% Ellis work summary %%%%

We aim to replicate the computational result presented in Figure S4 of their supplementary 
material (NOT SURE HOW TO CITE THIS) that correspond to the black curve on their plot. In this 
figure they show simulation results for the resonance spectral position of the lower frequency 
mode when having parallel polarization ($E^{\parallel}_{100}$), with an incidence angle of 22$^\circ$.
We first attempted to replicate this result since they have results where the gap between pillars 
is 5000 nm (10 times more than in the experiments and other simulations on the paper), and 
this diminishes the effects of coupling. This result is suitable to compare with 
what we can model using \pygbe.

\textbf{Difference between methods, mesh and dielectric data}

\textit{methods}
(I DID NOT MENTION ANYTHING RELATED TO WHAT THEY DO IN THE EXPERIMENTS, SHOULD WE INCLUDE THIS?)

In Ellis et al. simulations they solve full Maxwell equations via the RF package of the finite
element method software COMSOL. To represent the array of pillars and their interactions, they use
periodic boundary conditions of one pillar over a portion of substrate, and use periodic boundary conditions on it.
They present the reflectance as a function of the wave number. 
As we mentioned before, we use the boundary element method in 
the quasistatic approximation, which is suitable in these case since Ellis pillar's size is small 
compared to the wavelength involved in the simulations (800-1000 cm$^-1$ which is 10000-12500 nm). We measure 
extinction cross section, which will express as peaks instead of depressions (as seen in Ellis reflection plots)
since you see maximum extinction when there is minimum reflection. The intensity of the peaks is not comparable, but we are 
looking to match the wave number (variable of interest) at which these events happen.

\textit{mesh}
For the case of aspect ratio $AR=4$ we count with a triangular surface mesh provided by the authors of 
Ellis et al. Which is a non-structured (HOW IS IT CALLED WHEN IS FINER WHER THERE ARE CORNERS?) mesh of
approximately four thousand elements. To produce the remaining 
meshes, we use our python script to generate a uniform mesh and round the edges using Trimesh. 
After doing certain refinement studies we noticed that to achieve close results to the original mesh, using a
uniform mesh, we needed approximately double the density (I HAVE SOME PLOTS ON THIS RESULTS BUT 
I'M NOT SURE IF WE WANT TO PUT THEM).
To replicate the results of figure S4 on their supplementary material, we use uniform meshes. While for the 
replication of figure 2a and the validation result we used the mesh provided by the authors of Ellis et al.

\textit{dielectric data}
The dielectric data for the simulations, was given by the authors of the paper via a private communication, 
and corresponds to experimental values of the dielectric.

To be able to replicate the result of Figure S4 of the suplementary material of Ellis et al, we
need to identify the lower frequency mode for each of
the aspect ratios in our computations. For each aspect ratio (AR) 
value from 1 to 7 we computed the extinction cross section $C_{ext}$ across the wave number in the range
800-1000 $cm^{-1}$. We identify the lower frequency mode ($E^{\parallel}_{100}$ in Ellis paper)  that is not a 
longitudinal mode (mode related to the hight of the pillar, only visible when the incidence of the 
electric field is not normal). We can identify the longitudinal modes by comparing normal vs 22 degrees 
angle of incidence runs (see Figure \ref{fig:ellis_ang_inc}). As we mentioned the longitudinal modes, due to its nature, do not appear when we 
have the normal incidence . Then, on the 22 degree incidence computations we choose the lower wave number mode
that it's not a longitudinal mode.

Note: Ellis et al., changes the angle of incidence of the illuminating vector while in \pygbe we achieve this by
rotating the geometry instead.

\begin{figure}
    \centering
    \includegraphics[width=0.45\textwidth]{ellis_ang_inc.pdf} 
    \caption{Diagram of angles of incidence in \pygbe setup to comply with Ellis et al. disposition.}
    \label{fig:ellis_ang_inc}
 \end{figure}

Figure \ref{fig:AR_22_vs_norm} show the results of the extinction cross section of a SiC pillar of fixed
height ($H=950$ nm), fixed width ($W=400$) and varied length ($L=400-2800$ nm). The simulations where performed 
performed for the Long Edge orientation, meaning that the electric field is aligned with the length pillar 
when having normal incidence (see Figure \ref{fig:ellis_ang_inc}).

In Figure \ref{fig:AR_22_vs_norm} and Table \ref{tab:ar_peaks} we show the resonance peaks and
their corresponding wavelengths. In the table we mark in bold the peaks which correspond to 
the $E^{\parallel}_{100}$ mode.

\begin{figure}
    \centering
    \includegraphics[width=0.8\textwidth]{AR_22_vs_norm.pdf} 
    \caption{Extinction cross section across wave number for SiC pillars of multiple aspect ratios,  
             (H=950 nm, W=400nm, L=400-2800 nm (AR=1-7)), where we have normal incidence and a 
             22 degrees incidence.
            }
    \label{fig:AR_22_vs_norm}
 \end{figure}

\begin{table}
    \begin{center}
      \caption{Wavelength at which peaks happen for different aspect ratios, for runs where the electric
      field is parallel to the length (L) of the pillar. We have normal incidence and 22 degrees.}
      \label{tab:ar_peaks}
      \begin{tabular}{c c c c c c c c}
        \textbf{AR} \\
        \hline
        \multirow{2}{*}{1} & $\perp$ & \textbf{917.73} & 934.092 & 949.604 & 957.325 \\ % <-- Combining 2 rows with arbitrary with (*) and content 12
        & 22$^{\circ}$ & 883.926 & \textbf{917.73} & 935.052 & 949.604 & 957.325 \\ % <-- Content of first column omitted.
        \hline
        \multirow{2}{*}{2} & $\perp$ & \textbf{903.233} & 926.395 & 944.762 & 958.242 \\ % <-- Combining 2 rows with arbitrary with (*) and content 12
        & 22$^{\circ}$ & 896.517 & \textbf{903.233} & 926.395 & 944.762 & 958.242 \\ % <-- Content of first column omitted.
        \hline
        \multirow{2}{*}{3} & $\perp$ & \textbf{888.793} & 922.552 & 931.223 & 948.613 & 958.242 \\ % <-- Combining 2 rows with arbitrary with (*) and content 12
        & 22$^{\circ}$ & \textbf{888.793} & 899.418 & 922.552 & 931.223 & 958.242 \\ % <-- Content of first column omitted.
        \hline
        \multirow{2}{*}{4} & $\perp$ & \textbf{876.186} & 929.32 & 946.639 & 958.242 \\ % <-- Combining 2 rows with arbitrary with (*) and content 12
        & 22$^{\circ}$ & \textbf{876.186} & 901.281 & 921.618 & 929.32 & 945.745 & 958.242 \\ % <-- Content of first column omitted.
        \hline
        \multirow{2}{*}{5} & $\perp$ & \textbf{865.576} & 926.395 & 945.745 & 958.242 \\ % <-- Combining 2 rows with arbitrary with (*) and content 12
        & 22$^{\circ}$ & \textbf{865.576} & 901.281 & 921.618 & 958.242 \\ % <-- Content of first column omitted.
        \hline
        \multirow{2}{*}{6} & $\perp$ &  \textbf{856.904} & 914.793 & 923.489 & 929.32 & 946.639 & 958.242\\ % <-- Combining 2 rows with arbitrary with (*) and content 12
        & 22$^{\circ}$ & \textbf{856.904} & 901.281 & 920.6 & 958.242\\ % <-- Content of first column omitted.
        \hline
        \multirow{2}{*}{7} & $\perp$ &  \textbf{850.134} & 910.963 & 921.618 & 928.372 & 946.639 & 958.242 \\ % <-- Combining 2 rows with arbitrary with (*) and content 12
        & 22$^{\circ}$ & \textbf{850.134} & 901.281 & 910.963 & 920.6 & 958.242\\ % <-- Content of first column omitted.
        \hline
      \end{tabular}
    \end{center}
  \end{table}

Once we identify the proper modes, we proceed to the replication of Figure S4 on the 
supplementary material of Ellis et al work. It is worth mentioning that we limit ourselves to
the replication of the black curve in their plot, since in this case the pillars are separated 
by 5000 nm which means the coupling effects are negligible. Figure \ref{fig:rep_FS4_ellis} shows
the results from  Ellis et al (digitized using the WebPlotDigitizer) and the results obtained
with \pygbe, and Table \ref{tab:err_AR} shows that the percentage error is below 2$\%$ for all
the cases

\begin{figure}
    \centering
    \includegraphics[width=0.85\textwidth]{AR_rep_FS4_Ellis2016.pdf} 
    \caption{Replication of figure S4 of supplementary material of Ellis 2016. Wave
    number at which happens the $E^{\parallel}_{100}$ mode for different aspect ratios.}
    \label{fig:rep_FS4_ellis}
 \end{figure}

 \begin{table}
    \centering
    \caption{Percentage error for different aspect ratios.} 
    \label{tab:err_AR}
    \begin{tabular}{c c}
    \hline%\toprule
    AR & \% error \\
    \hline%\midrule
     $1$ & $0.95$ \\
     $2$ & $0.67$ \\
     $3$ & $0.35$ \\
     $4$ & $0.16$ \\
     $5$ & $0.72$ \\
     $6$ & $1.20$ \\
     $7$ & $1.59$ \\
    \hline%\bottomrule
    \end{tabular}
\end{table}

\subsubsection{Validation of \pygbe against Fig 2a experimental results}

For the case of aspect ratio four (AR=4) we obtain the smallest error using the uniform
mesh. Since we count with the mesh provided by the authors, and knowing that our computation 
for the mode $E^{\parallel}_{100}$ compares well with their computations, we attempted to 
validate our simulations with their experimental results (red curve on paper), as well 
as to replicate their simulations (green curve on paper) on Figure 2a of their paper.
Figure 2a of Ellis et al, presents measured and simulated reflectance of SiC pillar
arrays with a 500 nm gap. All their measurements and simulations were performed with 
22$^\circ$ off-normal angle of incidence and incoming polarization parallel to the 
elongated size of the pillar.

Using \pygbe we compute the extinction cross section of an isolated SiC pillar (AR=4)
with no substrate, submerged in air under a constant electric field in the z-direction, 
and rotate the orientation of the pillar to match the angle of incidence (see Figure \ref{fig:ellis_ang_inc}).
In Figure \ref{fig:pygbe_vs_exp_2a} we present 
comparison of our simulations and the experimental results from Ellis et al. There is a 
difference in the peaks wavelength that is noticeable. This could be attributed to the
fact that in their experiments the separation between pillars is of 500 nm, which implies
there are coupling effects that in our simulations are not contemplated.

\begin{figure}
    \centering
    \includegraphics[width=0.85\textwidth]{pygbe_vs_exp_fig2a_Ellis.pdf} 
    \caption{\pygbe vs experiments of figure 2a of Ellis 2016. Ellis data 
    was digitized with web digitizer.}
    \label{fig:pygbe_vs_exp_2a}
 \end{figure}

\paragraph{First order correction}

Since our simulations do not count for coupling effects, we can not strictly imitate 
the conditions to validate our solver. However, from Figure S4 on the supplementary 
material of Ellis et al, we know that coupling effects affect the $E^{\parallel}_{100}$ 
mode by a shift of 12.17 cm$^-1$. Therefore as a \textit{first order correction} we can 
subtract this value from our simulations to count for coupling effects. In Figure 
\ref{fig:val_2a} we present the result after applying the correction for coupling effects.
It is worth mentioning that the far left (837 cm$^-1$) and far right (964 cm$^-1$) peaks 
on Ellis results, are peaks associated with zone-folded LO (longitudinal) phonons, in 
particular P1 (837 cm$^-1$) is not infrared active \cite{ellis2016}. Ellis and coworkers, 
do not discuss this result into detail, but they concentrate their analysis on the peaks
that happen between 864–961 cm$^-1$

\begin{figure}
    \centering
    \includegraphics[width=0.85\textwidth]{validation_FOA_fig2a_Ellis.pdf} 
    \caption{Validation against experiments in figure 2a of Ellis 2016, using first order approximation}
    \label{fig:val_2a}
 \end{figure}

If we use the same approximation and compare the results with Ellis simulations on
Figure 2 of their paper (green curve) we get (see Figure \ref{fig:rep_2a})

\begin{figure}
    \centering
    \includegraphics[width=0.85\textwidth]{replication_FOA_fig2a_Ellis.pdf} 
    \caption{Replication of simulations in figure 2a of Ellis 2016, using first
     order approximation}
    \label{fig:rep_2a}
 \end{figure}